### 2023-10-07 R/community-clustering.qmd
initial commit 🚀

---
title: "Supplement A: Community Clustering"
date: today
published-title: "COMPILED"
format:
  html: 
    embed-resources: true
    fig-align: center
    fig-dpi: 300
    fig-responsive: true
    number-sections: true
    number-depth: 3
    toc: true
    toc-depth: 3
    smooth-scroll: true
    theme: cosmo
    highlight-style: a11y
execute: 
  echo: true
  warning: false
  error: false
---

## Introudction

**Goal:** identify geographic communities

**Data:** VEPII and Cibola databases

**Method:** using a somewhat novel clustering algorithm outlined below

This document is setup so that it focuses on the actual steps of the analysis, using verbose names for functions that should make their purpose clear. You can also dig into what each function does by unfolding the relevant code chunk. The critical R package for doing path analysis is [`cppRouting`](https://github.com/vlarmet/cppRouting). It's optimized for road networks, but still very fast. Someone should write a path-finding package (maybe using Rust?) optimized for grid networks...

## R Preamble

```{r}

library(arrow)
library(cppRouting)
library(ggfx)
library(ggnewscale)
library(ggspatial)
library(gt)
library(here)
library(igraph)
library(magick)
library(patchwork)
library(sf)
library(terra)
library(tidyterra)
library(tidyverse)

```

Some ggplot defaults

```{r}
#| code-fold: true

region_colors <- c(
  "cib" = "#D57300", 
  "cmv" = "#009E8C", 
  "nrg" = "#9B70FF"
)

region_labels <- as_labeller(c(
  "cmv" = "Central Mesa Verde",
  "nrg" = "Northern Rio Grande",
  "cib" = "Cibola"
))

theme_set(theme_bw())

theme_update(
  axis.text = element_text(size = rel(0.75), color = "gray50"),
  axis.ticks = element_line(linewidth = 0.2, color = "gray85"),
  axis.title = element_text(size = rel(1)),
  legend.text = element_text(size = rel(0.9)),
  legend.title = element_text(size = rel(0.9)),
  panel.grid = element_blank(),
  plot.title = element_text(size = rel(1), margin = margin(b = 2)),
  strip.background = element_blank(),
  strip.text = element_text(size = rel(1))
)

# trim all white space around image, but then add a few white pixels back (dx, dy)
prepare_image <- function(x, dx = 2, dy = 30, color = "white") {

  img <- magick::image_read(path = x)

  img <- magick::image_trim(img)

  info <- magick::image_info(img)

  new_width <- info$width + dx
  new_height <- info$height + dy

  img <- magick::image_extent(
    img, 
    geometry_area(new_width, new_height), 
    color = color
  )

  magick::image_write(img, path = x)

}

```

## Data

```{r}
#| code-fold: true

build_region <- function(gpkg, region){

  region_query <- paste0("SELECT * FROM regions WHERE region = '", region, "'")
  r <- read_sf(gpkg, query = region_query)

  site_query <- paste0("SELECT * FROM sites WHERE region = '", region, "'")
  s <- read_sf(gpkg, query = site_query) |> st_filter(r)

  if ("communities" %in% st_layers(gpkg)$name){

    community_query <- paste0("SELECT * FROM communities WHERE region = '", region, "'")
    t <- read_sf(gpkg, query = community_query) |> st_filter(r)

  } else {

    t <- NULL

  }

  list(
    name = region,
    preferred_projection = ifelse(region == "nrg", 26913, 26912),
    region = r,
    sites = s,
    communities = t
  )

}

add_elevation <- function(x){

  dem_fn <- paste0("dem-", x$name, ".tif")
  x$dem <- here("data", dem_fn) |> rast()

  return(x)

}

# getters
get_farms <- function(x){

  x$sites |> 
    st_drop_geometry() |> 
    filter(center == 0)

}

get_centers <- function(x){

  x$sites |> 
    st_drop_geometry() |> 
    filter(center == 1)

}

```

```{r}

gpkg <- here("data", "community-centers.gpkg")

cmv <- build_region(gpkg, "cmv") |> add_elevation()
nrg <- build_region(gpkg, "nrg") |> add_elevation()
cib <- build_region(gpkg, "cib") |> add_elevation()

```

```{r}
#| code-fold: true

e <- function(x){

  tibble(
    region = x$name,
    min =  terra::global(x$dem, min, na.rm = TRUE)$min,
    max =  terra::global(x$dem, max, na.rm = TRUE)$max,
    mean = terra::global(x$dem, mean, na.rm = TRUE)$mean,
    sd =   terra::global(x$dem, sd, na.rm = TRUE)$sd
  )

}

bind_rows(
  e(cmv),
  e(nrg),
  e(cib)
) |> 
  gt() |> 
  tab_header("Elevation") |> 
  fmt_number(-region, decimals = 1) |> 
  opt_align_table_header("left")

```

```{r}
#| code-fold: true

plot_region <- function(x){

  hell <- function(x, n = 1000){

    slope <- terrain(x, "slope", unit = "radians")
    aspect <- terrain(x, "aspect", unit = "radians")

    hill <- shade(slope, aspect, 30, 45)
    hill <- setValues(hill, scales::rescale(values(hill), to = c(1, n)))
    hill <- round(hill)

    list(
      shade = hill,
      grays = hcl.colors(n, "Grays")[values(hill)]
    )

  }

  h <- hell(x$dem)

  ggplot() +
    geom_spatraster(
      data = h$shade, 
      fill = h$grays,
      maxcell = Inf
    ) +
    geom_spatraster(data = x$dem, alpha = 0.5) +
    scale_fill_hypso_c(
      name = "Elevation (m)",
      palette = "utah_1",
      limits = c(1350, 3850),
      alpha = 0.5,
      na.value = "transparent"
    ) +
    with_outer_glow(
      geom_sf(
        data = x$sites |> filter(center == 0),
        shape = 16,
        color = "gray25",
        alpha = 0.5,
        size = 0.6
      ),
      colour = "white",
      sigma = 0.5,
      expand = 1
    ) +
    with_outer_glow(
      geom_sf(
        data = x$sites |> filter(center == 1),
        shape = 17,
        color = "#CD3C00",
        alpha = 0.8,
        size = 1.1
      ),
      colour = "white",
      sigma = 0.5,
      expand = 1
    ) +
    annotation_scale(
      aes(location = "bl", line_col = "white", text_col = "white"),
      height = unit(0.18, "cm"),
      line_width = 0.5
    ) +
    coord_sf(expand = FALSE) +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      legend.position = "none",
      plot.margin = margin(l = 2)
    )

}

gg <- (plot_region(cmv) + ggtitle("Central Mesa Verde")) +
  (plot_region(nrg) + ggtitle("Northern Rio Grande")) + 
  (plot_region(cib) + ggtitle("Cibola"))

fn <- here("figures", "overview-maps.png")

ggsave(
  plot = gg,
  filename = fn,
  width = 7.5, 
  height = 4.5,
  dpi = 600,
  bg = "white"
)

prepare_image(fn)

remove(plot_region, gg, fn)

```

::: {#fig-overview}
![](../figures/overview-maps.png){width="100%" fig-alt="An overview map showing the distribution of farms and community centers across the three study areas."}

An overview map showing the distribution of farms and community centers across the three study areas.
:::

## Graph

It's important to note that throughout we track movement between grid cells, not site points.

### Convert grid to graph

```{r}
#| code-fold: true

add_graph <- function(x, max_slope = NULL, ...){

  g <- terra::adjacent(
    x$dem,
    cells = terra::cells(x$dem),
    directions = 8,
    pairs = TRUE
  )

  g <- as.data.frame(g)

  g <- g[!g$to %in% terra::cells(x$dem, NA_real_)[[1]], ]

  # distance between cells
  g$run <- terra::distance(
    terra::xyFromCell(x$dem, as.vector(g$from)),
    terra::xyFromCell(x$dem, as.vector(g$to)),
    lonlat = terra::is.lonlat(x$dem),
    pairwise = TRUE
  )

  # difference between cells
  g$from_elevation <- terra::extract(x$dem, as.vector(g$from))[,1]

  g$to_elevation <- terra::extract(x$dem, as.vector(g$to))[,1]

  g$rise <- g$from_elevation - g$to_elevation

  # slope
  g$slope <- g$rise/g$run

  # what is a realistic slope people would try to hike?
  if (!is.null(max_slope)) {

    g <- g[g$slope <= tan(max_slope * pi / 180), ]

  }

  g$speed <- campbell(g$slope, ...)

  g$cost <- g$run/g$speed

  g <- g[, c("from", "to", "cost")]

  coords <- cbind(
    ID = terra::cells(x$dem), 
    terra::xyFromCell(x$dem, terra::cells(x$dem))
  ) |> 
    as.data.frame() |> 
    rename_with(toupper)

  # force as.character(1000000) to be "1000000" instead of "1e-6"
  # this fixes problem with makegraph()
  options(scipen = 9999)

  x$graph <- makegraph(
    as.data.frame(g),
    directed = TRUE,
    coords = coords
  )

  return(x)

}

campbell <- function(x, decile = 30L) {

  if (length(decile) != 1) {

    stop("campbell accepts only one decile at a time.")

  }

  # campbell function wants degrees, not rise-over-run as in tobler
  # but, we're working with rr, so remember to convert radians to degrees with 180/pi!
  slope <- atan(x) * 180 / pi

  # values provided in Campbell 2019, Supplement, Table S1
  # for simplicity, I excluded all but the deciles
  deciles <-
    data.frame(
      decile = c(10L, 20L, 30L, 40L, 50L, 60L, 70L, 80L, 90L),
      a = c(-1.568, -1.71,  -1.858, -1.958, -2.171, -2.459, -2.823,  -3.371,  -3.06),
      b = c(13.328, 10.154,  8.412,  8.96,  10.064, 11.311, 12.784,  15.395,  16.653),
      c = c(38.892, 36.905, 39.994, 50.34,  63.66,  79.287, 98.697, 134.409, 138.875),
      d = c( 0.404,  0.557,  0.645,  0.649,  0.628,  0.599,  0.566,   0.443,   0.823),
      e = c(-0.003, -0.004, -0.004, -0.005, -0.005, -0.005, -0.005,  -0.005,  -0.0139)
    )

  ind <- which(deciles$decile == as.integer(decile))

  deciles <- deciles[ind, ]

  a <- deciles$a
  b <- deciles$b
  c <- deciles$c
  d <- deciles$d
  e <- deciles$e

  # lorentz distribution
  lorentz <- (1 / ((pi * b) * (1 + ((slope - a) / b)^2)))

  # modified lorentz
  (c * lorentz) + d + (e * slope)

}

```

```{r}

cmv <- cmv |> add_graph(max_slope = 45, decile = 10L)
nrg <- nrg |> add_graph(max_slope = 45, decile = 10L)
cib <- cib |> add_graph(max_slope = 45, decile = 10L)

```

### Add distance matrix

```{r}
#| code-fold: true

add_distance_matrix <- function(x){

  x$sites$cell <- terra::cells(x$dem, terra::vect(x$sites))[,2]

  # use unique cell IDs as some centers are in the same pixel, 
  # so calculations would be redundant
  dm <- get_distance_matrix(
    x$graph,
    from = x$sites$cell |> unique(),  
    to = x$sites$cell |> unique()
  )

  # make symmetric by averaging off-diagonals
  d <- cbind(dm[lower.tri(dm)], t(dm)[lower.tri(dm)])

  dm[lower.tri(dm)] <- rowMeans(d, na.rm = TRUE)
  dm[upper.tri(dm)] <- t(dm)[upper.tri(dm)]

  diag(dm) <- 0

  x$dm <- dm

  return(x)

}

```

```{r}

cmv <- cmv |> add_distance_matrix()
nrg <- nrg |> add_distance_matrix()
cib <- cib |> add_distance_matrix()

```

```{r}
#| code-fold: true

get_commutes <- function(x, n){

  centers <- get_centers(x) |> select(-center)
  farms   <- get_farms(x) |> select(-center)

  # subset dm (distance matrix) to get dm[farms, centers]
  i <- which(rownames(x$dm) %in% farms$cell)
  j <- which(colnames(x$dm) %in% centers$cell)

  dm <- x$dm[i, j]

  observed <- apply(dm, 1, min)
  observed <- density(observed, kernel = "gaussian", n = 512, from = 0, to = 7200)

  r <- terra::rast(
    nrows = nrow(x$dem), 
    ncols = ncol(x$dem), 
    extent = ext(x$dem), 
    crs = crs(x$dem)
  )

  dmax <- st_distance(
    x$sites |> filter(center == 0), 
    x$sites |> filter(center == 1)
  ) |> units::drop_units() |> apply(1, min) |> max()

  i <- cells(
    x$dem, 
    x$sites |> filter(center == 1) |> st_buffer(dmax) |> st_union() |> vect()
  )[, "cell"]

  i <- i[i %in% x$graph$dict$ref]

  r[i] <- 1

  # don't sample cells with centers in them
  j <- colnames(dm) |> as.numeric()

  r[j] <- NA

  random_points <- terra::spatSample(
    r, 
    size = n,
    method = "regular",
    cells = TRUE,
    na.rm = TRUE
  ) |> pull(cell) |> unique()

  dm <- get_distance_matrix(
    x$graph,
    from = random_points,  
    to = colnames(dm)
  )

  # make symmetric by averaging off-diagonals
  d <- cbind(dm[lower.tri(dm)], t(dm)[lower.tri(dm)])

  dm[lower.tri(dm)] <- rowMeans(d, na.rm = TRUE)
  dm[upper.tri(dm)] <- t(dm)[upper.tri(dm)]

  diag(dm) <- 0

  expected <- apply(dm, 1, min)
  expected <- density(expected, kernel = "gaussian", n = 512, from = 0, to = 7200)

  tibble(
    region = x$name,
    distance = observed$x,
    observed = observed$y,
    expected = expected$y
  )

}

set.seed(1701) # USS Enterprise

commutes <- bind_rows(
  cmv |> get_commutes(n = 1000),
  nrg |> get_commutes(n = 1000),
  cib |> get_commutes(n = 1000)
)

q <- function(x, p){

  centers <- get_centers(x) |> select(-center)
  farms   <- get_farms(x) |> select(-center)

  # subset dm (distance matrix) to get dm[farms, centers]
  i <- which(rownames(x$dm) %in% farms$cell)
  j <- which(colnames(x$dm) %in% centers$cell)

  dm <- x$dm[i, j]

  observed <- apply(dm, 1, min)

  tibble(
    region = x$name,
    distance = quantile(observed, p = p) |> unname()
  )

}

q95 <- bind_rows(
  q(cmv, 0.95),
  q(nrg, 0.95),
  q(cib, 0.95)
) |> 
  mutate(
    y = c(2.4, 1.8, 1.2),
    label = case_when(
      region == "cib" ~ "Cibola",
      region == "cmv" ~ "Central Mesa Verde",
      region == "nrg" ~ "Northern Rio Grande",
      TRUE ~ NA
    )
  )

gg <- ggplot() +
  geom_hline(
    yintercept = 0,
    color = "gray75",
    linewidth = 0.2
  ) +
  geom_line(
    data = commutes, 
    aes(distance/60, log(observed/expected), color = region), 
    linewidth = 0.9,
    lineend = "round"
  ) +
  geom_point(
    data = q95,
    aes(distance/60, y, color = region),
    size = 2
  ) +
  geom_text(
    data = q95,
    aes(distance/60, y, color = region, label = label),
    vjust = 0.5,
    hjust = 0,
    nudge_x = 2,
    size = 12/.pt
  ) +
  scale_color_manual(name = NULL, values = region_colors) +
  annotate(
    "point",
    x = min(q95$distance)/60,
    y = 3.0,
    color = "gray25",
    size = 2
  ) +
  annotate(
    "text",
    label = "95% Quantile of Observed",
    x = min(q95$distance)/60 + 2,
    y = 3.0,
    color = "gray25",
    vjust = 0.5,
    hjust = 0,
    size = 12/.pt
  ) +
  labs(
    x = "Commute to Nearest Center [mins]",
    y = "Density\nlog(Observed) - log(Random)"
  ) +
  scale_x_continuous(breaks = seq(0, 120, by = 30), expand = expansion(0.03)) +
  coord_cartesian(xlim = c(0, 120)) +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_line(linewidth = 0.2, color = "gray85")
  )

fn <- here("figures", "commute-kl.png")

ggsave(
  plot = gg,
  filename = fn,
  width = 5,
  height = 3.3,
  dpi = 600
)

prepare_image(fn)

remove(get_commutes, q, q95, gg)

```

::: {#fig-commutes}
![](../figures/commute-kl.png){width="60%" fig-alt="Difference between log probability density of obversed commute times from farms to nearest centers and log probability density of commute times from random locations to nearest centers."}

Difference between log probability density of obversed commute times from farms to nearest centers and log probability density of commute times from random locations to nearest centers.
:::

## Communities

Our community detection algorithm proceeds as follows:

1. Identify community centers   
2. Identify farms within commute distance 
3. Join communities to their overlapping neighbors 
4. Exclude farms closer to another community center (tie-breaking step) 
5. Draw smallest concave hull encompassing all farms, centers, and paths

Step 1 is already done and stored in the data.

### Filter by commute distance

```{r}
#| code-fold: true

subset_distance_matrix <- function(x){

  centers <- get_centers(x) |> select(-center)
  farms   <- get_farms(x) |> select(-center)

  # subset dm (distance matrix) to get dm[farms, centers]
  i <- which(rownames(x$dm) %in% farms$cell)
  j <- which(colnames(x$dm) %in% centers$cell)

  x$dm[i, j]

}

filter_distance_matrix <- function(x, d){

  nearest_farms <- apply(x, 1, \(x){ which(x <= d) })
  nearest_farms <- lapply(nearest_farms, \(x){ as.numeric(names(x)) })

  arrow_table(
    center = unlist(nearest_farms),
    farm = rep(names(nearest_farms), lengths(nearest_farms)) |> as.numeric()
  ) |> arrange(center, farm)

}

find_nearby_farms <- function(x, dmax){

  x$dmax <- dmax

  x$nearest_farms <- x |> 
    subset_distance_matrix() |> 
    filter_distance_matrix(dmax)

  sites <- x$sites |> 
    st_drop_geometry() |> 
    select(cell, n_room) |> 
    group_by(cell) |> 
    summarize(n_room = sum(n_room))

  x$nearest_farms <- x$nearest_farms |> left_join(sites, by = join_by(farm == cell))

  return(x)

}

```

```{r}

# how many seconds in ... ?
one_hour <- 3600

cmv <- cmv |> find_nearby_farms(dmax = one_hour)
nrg <- nrg |> find_nearby_farms(dmax = one_hour)
cib <- cib |> find_nearby_farms(dmax = one_hour)

```

### Find neighboring centers

> For any two centers $C_1$ and $C_2$ with nearby farm room counts $N_1 < N_2$, if $pN_1$ of $C_1$’s rooms are within a distance $d$ of $C_1$ and $C_2$, then $C_1$ is part of the same community as $C_2$.

```{r}
#| code-fold: true

join_neighboring_centers <- function(x, p, djoin){

  x$djoin <- djoin

  nearest_farms <- x |> 
    subset_distance_matrix() |> 
    filter_distance_matrix(djoin)

  sites <- x$sites |> 
    st_drop_geometry() |> 
    select(cell, n_room) |> 
    group_by(cell) |> 
    summarize(n_room = sum(n_room))

  nearest_farms <- nearest_farms |> left_join(sites, by = join_by(farm == cell))

  remove(sites)

  room_counts <- nearest_farms |> 
    group_by(center) |> 
    summarize(n_room = sum(n_room)) |> 
    collect()

  R <- outer(room_counts$n_room, room_counts$n_room, FUN = '<=')
  diag(R) <- FALSE

  colnames(R) <- room_counts$center
  rownames(R) <- room_counts$center

  R <- apply(R, 1, which)
  R <- lapply(R, \(x){ as.numeric(names(x)) })

  neighbors_matrix <- imap(R, \(x, idx){

    N <- room_counts |> filter(center == as.numeric(idx)) |> pull(n_room)

    focal_farms <- nearest_farms |> 
      filter(center == as.numeric(idx)) |> 
      collect() |> 
      pull(farm)

    cntrs <- nearest_farms |> 
      filter( # this filter gives the intersection of farms/cells 
        center %in% x,
        farm %in% focal_farms
      ) |> 
      group_by(center) |> 
      summarize(proportion = sum(n_room)/N) |> 
      filter(proportion >= p) |> 
      collect() |> 
      pull(center)

    names(R) %in% cntrs

  })

  neighbors_matrix <- do.call("rbind", neighbors_matrix)
  diag(neighbors_matrix) <- TRUE

  colnames(neighbors_matrix) <- names(R)
  rownames(neighbors_matrix) <- names(R)

  # igraph...
  g <- igraph::graph.adjacency(neighbors_matrix)
  membership <- igraph::components(g)$membership

  x$lookup <- tibble(
    center = names(membership) |> as.numeric(),
    community = unname(membership)
  )

  return(x)

}

```

```{r}

# how many seconds in ... ?
half_hour <- 1800

cmv <- cmv |> join_neighboring_centers(p = 0.8, djoin = half_hour)
nrg <- nrg |> join_neighboring_centers(p = 0.8, djoin = half_hour)
cib <- cib |> join_neighboring_centers(p = 0.8, djoin = half_hour)

```

### Find nearest center

```{r}
#| code-fold: true

find_nearest_center <- function(x){

  farms_to_centers <- subset_distance_matrix(x)

  # find the nearest center for each farm
  nearest_center <- apply(farms_to_centers, 1, \(x) which.min(x))
  nearest_center <- colnames(farms_to_centers)[nearest_center]

  distance <- apply(farms_to_centers, 1, min)

  x$nearest_center <- tibble(
    farm = farms_to_centers |> rownames() |> as.numeric(),
    center = nearest_center |> as.numeric(),
    distance = unname(distance)
  ) |> filter(distance <= x$dmax)

  return(x)

}

```

```{r}

cmv <- cmv |> find_nearest_center() 
nrg <- nrg |> find_nearest_center() 
cib <- cib |> find_nearest_center()

```

### Define boundaries

Now that we have each farm associated with its nearest community center, and each center joined with its neighbors, assuming it has any, we can define a boundary for the community using a concave hull. We choose a concave rather than convex hull because the latter tends to artificially inflate the size of the community, often to a significant degree. On the other hand, the concave hull can generate extremely unrealistic shapes for communities. To avoid these issues, we first compute the least-cost path between each farm in the community and every other farm in the community. We then add the vertices that define these paths as virtual points to the community. We then generate the concave hull for the expanded set of community points. The consequence is that an individual setting off from one farm in the community to another should never have to leave the community. Their walk will always be within the community boundary.

The result is still not perfect, however, so we further simplify polygons (including filling holes) by adding then removing a small buffer. Two things to note here: (1) we project the boundaries first as Google's `S2` system for geocomputations on the sphere (the default now in `{sf}`) is janky with how it does buffering and (2) we add a small buffer to smooth some of the noise in the hull.

```{r}
#| code-fold: true

collect_members <- function(x){

  lookup <- x$lookup

  centers <- get_centers(x) |> 
    left_join(lookup, by = join_by(cell == center)) |> 
    select(site_id, n_room, initial_phase, final_phase, community) |> 
    filter(!is.na(community)) |> 
    mutate(type = "center")

  farms <- get_farms(x) |> 
    select(-center) |> 
    left_join(x$nearest_center, by = join_by(cell == farm)) |> 
    left_join(lookup, by = "center") |> 
    select(site_id, n_room, initial_phase, final_phase, community) |> 
    filter(!is.na(community)) |> 
    mutate(type = "farm")

  x$members <- bind_rows(farms, centers) |> 
    arrange(community, site_id) |> 
    left_join(x$sites[, "site_id"], by = "site_id") |> 
    mutate(region = x$name, .before = everything()) |> 
    st_sf()

  return(x)

}

define_boundaries <- function(x, concavity_ratio, dilate){

  community_ids <- x$members |> pull(community) |> unique()

  communities <- vector("list", length(community_ids))

  # cli::cli_progress_bar(
  #   "Define boundaries", 
  #   total = length(community_ids)
  # )

  for (i in community_ids){

    m <- x$members |> filter(community == i)

    cells <- terra::cells(x$dem, vect(m))[, 2] |> unique()

    paths <- get_multi_paths(
      x$graph, 
      from = cells, 
      to = cells
    )

    paths <- paths |> unlist(recursive = TRUE) |> as.numeric() |> unique()

    xy <- rbind(
      terra::xyFromCell(x$dem, paths),
      st_coordinates(m)[, 1:2]
    )

    xy <- xy |> st_multipoint() |> st_sfc(crs = st_crs(m)) |> st_cast("POINT")

    m <- st_geometry(m)

    border <- st_sf(geometry = c(xy, m)) |> 
      summarize() |> 
      st_concave_hull(ratio = concavity_ratio) |> 
      mutate(community = i)

    communities[[i]] <- border

    remove(i, m, cells, paths, xy, border)

    # cli::cli_progress_update()

  }

  x$communities <- communities |> 
    bind_rows() |> 
    st_transform(x$preferred_projection) |> 
    st_buffer(dilate[1]) |> 
    st_buffer(dilate[2]) |> 
    st_transform(4326) |> 
    mutate(region = x$name, .before = "community")

  return(x)

}

```

```{r}

cmv <- cmv |> 
  collect_members() |> 
  define_boundaries(concavity_ratio = 0.1, dilate = c(130, -100))

nrg <- nrg |> 
  collect_members() |> 
  define_boundaries(concavity_ratio = 0.1, dilate = c(130, -100))

cib <- cib |> 
  collect_members() |> 
  define_boundaries(concavity_ratio = 0.1, dilate = c(130, -100))

```

### Drop small communities

```{r}
#| code-fold: true

drop_small_communities <- function(x, threshold){

  final_list <- x$members |> 
    st_drop_geometry() |> 
    filter(type == "farm") |> 
    group_by(community) |> 
    summarize(n = n()) |> 
    filter(n > threshold) |> 
    pull(community)

  x$communities <- x$communities |> filter(community %in% final_list)

  x$members <- x$members |> st_filter(x$communities)

  return(x)

}

```

```{r}

cmv <- cmv |> drop_small_communities(threshold = 3)
nrg <- nrg |> drop_small_communities(threshold = 3)
cib <- cib |> drop_small_communities(threshold = 3)

```

## Results

Here we show the community polygons themselves as well as the log-density of rooms in each.

```{r}
#| code-fold: true

visualize_communities <- function(x, lbl = region_labels){

  hell <- function(x, n = 1000){

    slope <- terrain(x, "slope", unit = "radians")
    aspect <- terrain(x, "aspect", unit = "radians")

    hill <- shade(slope, aspect, 30, 45)
    hill <- setValues(hill, scales::rescale(values(hill), to = c(1, n)))
    hill <- round(hill)

    list(
      shade = hill,
      grays = hcl.colors(n, "Grays")[values(hill)]
    )

  }

  h <- hell(x$dem)

  members <- x$members |> 
    st_drop_geometry() |> 
    group_by(community) |> 
    summarize(n_room = sum(n_room))

  communities <- x$communities |> 
    left_join(members, by = "community") |> 
    mutate(
      area = st_area(geometry) |> units::set_units(km2) |> units::drop_units(),
      density = n_room/area,
      density = log(density+1),
      density = cut(density, 9, labels = FALSE)
    )

  ggplot() +
    ggtitle(lbl(x$name)[[1]]) +
    geom_spatraster(
      data = h$shade, 
      fill = h$grays,
      maxcell = Inf
    ) +
    geom_spatraster(data = x$dem, alpha = 0.5,) +
    scale_fill_distiller(palette = "Greys", guide = "none") +
    ggnewscale::new_scale_fill() +
    geom_sf(
      data = communities,
      fill = "transparent",
      color = "white",
      alpha = 0.1,
      linewidth = 0.5
    ) +
    geom_sf(
      data = communities,
      aes(fill = density, color = density),
      linewidth = 0.1
    ) + 
    scale_color_gradientn(
      name = "Log Density", 
      colors = RColorBrewer::brewer.pal(9, "YlOrBr"),
      breaks = c(1, 5, 9),
      labels = c("Low", "Medium", "High"),
      guide = guide_legend()
    ) +
    scale_fill_gradientn(
      name = "Log Density", 
      colors = RColorBrewer::brewer.pal(9, "YlOrBr") |> alpha(0.6),
      breaks = c(1, 5, 9),
      labels = c("Low", "Medium", "High"),
      guide = guide_legend()
    ) +
    annotation_scale(
      aes(location = "bl", line_col = "white", text_col = "white"),
      height = unit(0.18, "cm"),
      line_width = 0.5
    ) +
    coord_sf(expand = FALSE) +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      legend.key = element_rect(color = "gray30", fill = "transparent"),
      plot.margin = margin(l = 2)
    )

}

zg <- list(cmv, nrg, cib) |> 
  map(visualize_communities) |> 
  patchwork::wrap_plots(ncol = 3)

# patchwork is great, but not perfect...
bob <- theme(
  legend.box.margin = margin(t = -10),
  legend.key.size = unit(0.5, "cm"),
  legend.margin = margin(),
  legend.justification = "left",
  legend.position = "bottom"
)

# had to fiddle with this to get the sizes consistent with the overview maps
fn <- here("figures", "communities-all.png")

ggsave(
  plot = zg + plot_layout(guides = "collect") & bob,
  filename = fn,
  width = 7.5, 
  height = 4.5,
  dpi = 600,
  bg = "white"
)

prepare_image(fn)

remove(visualize_communities, zg, bob, fn)

```

::: {#fig-communities}
![](../figures/communities-all.png){width="100%" fig-alt="Map shows defined community boundaries with fill colors representing the log density of rooms in each."}

Map shows defined community boundaries with fill colors representing the log density of rooms in each.
:::

### Community Distributions

```{r}
#| code-fold: true

get_community_data <- function(x){

  # AREA
  A <- x$communities |> 
    st_area() |> 
    units::set_units(km2) |> 
    units::drop_units()

  tbl_area <- tibble(
    region = x$name, 
    community = x$communities$community,
    area = A
  )

  # ROOMS
  tbl_rooms <- x$sites |> 
    st_join(x$communities[, "community"]) |> 
    select(region, community, n_room) |> 
    st_drop_geometry() |> 
    group_by(community) |> 
    summarize(
      region = unique(region),
      n_room = sum(n_room)
    ) |> 
    relocate(region)

  # FARMS
  tbl_sites <- x$members |> 
    st_drop_geometry() |> 
    filter(type == "farm") |> 
    group_by(community) |> 
    summarize(n_farm = n(), region = unique(region))

  # CENTERS
  tbl_centers <- x$members |> 
    st_drop_geometry() |> 
    filter(type == "center") |> 
    group_by(community) |> 
    summarize(n_center = n(), region = unique(region))

  # COMMUTES
  cell_lookup <- x$sites |> 
    st_drop_geometry() |> 
    select(site_id, cell)

  tbl_commutes <- x$members |> 
    st_drop_geometry() |> 
    filter(type == "center") |> 
    left_join(cell_lookup, by = "site_id") |> 
    left_join(x$nearest_center, by = join_by(cell == center), relationship = "many-to-many") |> 
    group_by(community) |> 
    summarize(commute = mean(distance, na.rm = TRUE), region = unique(region))

  # GATHER EVERYTHING INTO ONE TABLE
  tbl_area |> 
    left_join(tbl_rooms, by = c("region", "community")) |> 
    left_join(tbl_sites, by = c("region", "community")) |>
    left_join(tbl_centers, by = c("region", "community")) |> 
    left_join(tbl_commutes, by = c("region", "community")) |> 
    mutate(room_density = n_room/area)

}

```

```{r}

results <- list(cmv, nrg, cib) |> map(get_community_data) |> bind_rows()

```

```{r}
#| code-fold: true

results |> 
  mutate(commute = commute/60) |> 
  group_by(region) |> 
  summarize(across(area:room_density, list("η" = median, "µ" = mean, "σ" = sd))) |> 
  rename_with(str_remove, pattern = "n_|room_") |> 
  pivot_longer(
    !region,
    names_to = c("variable", "statistic"),
    names_sep = "_"
  ) |> 
  pivot_wider(
    names_from = "variable",
    values_from = "value"
  ) |> 
  mutate(across(area:density, \(x) paste0(statistic, ": ", round(x,2)))) |> 
  group_by(region) |> 
  summarize(across(area:density, \(x) paste0(x, collapse = "<br>"))) |> 
  rename(
    " " = region,
    "Area (km2)" = area,
    "Rooms (N)" = room,
    "Farms (N)" = farm,
    "Centers (N)" = center,
    "Commute (mins)" = commute,
    "Room Density (N/km2)" = density
  ) |> 
  slice(c(2, 3, 1)) |> 
  gt() |> 
  tab_header(
    title = "Community summaries",
    subtitle = md("η = median, µ = mean, σ = standard deviation")
  ) |> 
  fmt_markdown(columns = everything()) |> 
  cols_width(
    " " ~ pct(8),
    "Area (km2)" ~ pct(13),
    "Rooms (N)" ~ pct(13),
    "Farms (N)" ~ pct(13),
    "Centers (N)" ~ pct(13),
    "Commute (mins)" ~ pct(18),
    "Room Density (N/km2)" ~ pct(22)
  ) |> 
  opt_align_table_header("left")

```

```{r}
#| eval: false
#| echo: false

# this code is just for exploring communities
# and associating important (re: named) centers with them
plot_labels <- function(x){

  ggplot(x$communities) +
    geom_sf() +
    geom_sf_label(aes(label = community)) +
    theme_void()

}

get_members <- function(x){

  x$members |> 
    st_drop_geometry() |> 
    mutate(region = x$name, .before = everything())

}

get_sites <- function(x){ x$sites |> st_drop_geometry() }

results |> 
  group_by(region) |> 
  slice_max(room_density, n = 5) |> 
  left_join(
    map(list(cmv, nrg, cib), get_members) |> bind_rows(), 
    by = c("region", "community")
  ) |> 
  left_join(
    map(list(cmv, nrg, cib), get_sites) |> bind_rows() |> filter(center == 1), 
    by = c("region", "site_id")
  ) |> 
  select(region, community, site_id, site_name, area) |> 
  filter(region == "cib") |> View()

```

```{r}
#| code-fold: true

lbls <- results |> 
  group_by(region) |> 
  summarize(x = quantile(area/n_center, p = 0.75)) |> 
  mutate(
    variable = "center_area",
    y = c(1.25, 3.25, 2.25),
    label = case_when(
      region == "cib" ~ "Cibola",
      region == "cmv" ~ "Central Mesa Verde",
      region == "nrg" ~ "Northern Rio Grande",
      TRUE ~ region
    )
  )

lblr <- as_labeller(c(
  "commute" = "Mean Commute to Nearest Center [mins]",
  "room_density" = "Room Density [N/km2]",
  "farm_area" = "Mean Area for Farms [km2/N]",
  "center_area" = "Mean Area for Centers [km2/N]",
  "farm_ratio" = "Ratio of Farms [N] to Centers [N]",
  "area" = "Total Area [km2]"
))

gg <- results |> 
  mutate(
    farm_area = area/n_farm,
    center_area = area/n_center,
    commute = commute/60,
    farm_ratio = n_farm/n_center
  ) |> 
  select(-n_room, -n_farm, -n_center) |> 
  pivot_longer(c(-region, -community), names_to = "variable") |> 
  ggplot() + 
  geom_boxplot(aes(
    x = value, 
    y = fct_relevel(region, "cib", "nrg"), 
    fill = region
  )) +
  geom_text(
    data = lbls,
    aes(x, y, label = label, color = region),
    size = 11.5/.pt,
    nudge_x = 1,
    hjust = 0
  ) +
  scale_color_manual(values = region_colors) +
  scale_fill_manual(values = alpha(region_colors, 0.6)) +
  labs(
    x = NULL,
    y = NULL
  ) +
  facet_wrap(
    vars(variable), 
    ncol = 2, 
    nrow = 3,
    scale = "free_x",
    strip.position = "bottom",
    labeller = lblr
  ) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none",
    panel.grid.major.x = element_line(linewidth = 0.2, color = "gray85"),
    strip.placement = "outside",
    strip.text = element_text(size = rel(1), margin = margin(t = -1, b = 10))
  )

fn <- here("figures", "results-boxplot.png")

ggsave(
  plot = gg,
  filename = fn,
  width = 6.5,
  height = 7.0,
  dpi = 600
)

prepare_image(fn)

remove(lbls, lblr, gg, fn)

```

::: {#fig-distributions}
![](../figures/results-boxplot.png){width="75%" fig-alt="Distributions of important variables across communities and regions."}

Distributions of important variables across communities and regions.
:::

### Save results

```{r}

bind_rows(
  cmv$members |> mutate(region = "cmv", .before = site_id),
  nrg$members |> mutate(region = "nrg", .before = site_id),
  cib$members |> mutate(region = "cib", .before = site_id)
) |> 
  st_drop_geometry() |> 
  write_sf(gpkg, layer = "members")

bind_rows(
  cmv$communities, 
  nrg$communities, 
  cib$communities
) |> 
  left_join(results, by = c("region", "community")) |> 
  write_sf(gpkg, layer = "communities")

```

## Session Info

```{r}
#| code-fold: true

# save the session info as an object
pkg_sesh <- sessioninfo::session_info(pkgs = "attached")

# inject the quarto info
pkg_sesh$platform$quarto <- paste(
  quarto::quarto_version(), 
  "@", 
  quarto::quarto_path()
)

# print it out
pkg_sesh

```