### 2024-03-02 manuscript/manuscript.qmd
multiple changes based on rev reqs

---
title: A method for defining dispersed community territories
date: today
author:
  - name: Kenneth B. Vernon
    email: Kenneth.Vernon@colorado.edu
    affiliations: 
      - id: colorado
        address: | 
          Center for Collaborative Synthesis in Archaeology,  
          Institute of Behavioral Science,  
          University of Colorado, Boulder,  
          1440 15th Street,  
          Boulder, CO 80302
    attributes:
        corresponding: true 
  - name: Scott G. Ortman
    affiliations:
      - ref: colorado
abstract: |
  The transition from dispersed to aggregated forms of settlement reflects a critical shift in the relative value of social and primary (food) modes of production. However, investigating trade-offs between them requires estimates of the extent of community territories, including their nearby arable land. Here we demonstrate a simple algorithm to do that. Our algorithm is analogous to that used to define core-based statistical areas for the US census, though instead of modern metropolitan areas, we rely on community centers (or areas of known and persistent interaction between unrelated individuals). We provide examples of our algorithm by applying it to archaeological sites in the central Mesa Verde, northern Rio Grande, and Cibola regions. A sensitivity analysis is also conducted to demonstrate how each tuning parameter contributes to the algorithm.
keywords:
  - Aggregation
  - Core-Based Statistical Area
  - Community Detection
  - Complex Systems
  - Least-Cost Path Analysis
  - Population Density
target-journal: Journal of Archaeological Science
bibliography: bibliography.bib
corresponding_author: Kenneth.Vernon@colorado.edu
linenumbers: false
format: 
  simple-pdf: default
  docx: default
execute: 
  echo: false
  error: false
  # warning: false
engine: knitr
---

# Introduction

Urban geographers recognize density as a crucial property of human communities
which influences many other properties, including productivity, resource needs,
and infectious disease rates [@angel_etal2021; @bettencourt2021;
@duranton_puga2020]. The archaeological record of subsistence farming societies
contains physical traces of community formations with a wide range of densities,
often glossed as "dispersed" or "aggregated" [@birch2013; @drennan_etal2015;
@gyucha2019]. The primary difference between the two has to do with the spatial
distribution of residents relative to primary (food) production. In dispersed
formations, households are scattered across an area, interspersed with the land
that they farmed, thus requiring them to commute to central places for various
forms of social production. In aggregated formations, in contrast, households
are clustered in a village or town, such that they reside in locations of social
production, but must commute to fields for primary production.

From a complex systems perspective, the main way these two settlement
morphologies vary is with respect to which costs are being minimized. In the
dispersed pattern, costs associated with primary (food) production by the
household are being minimized, whereas in the aggregated pattern, it is costs
associated with various forms of social production, from government and ritual
to economic exchange and warfare. Presumably, choices regarding which sorts of
costs to minimize are driven by the relative productivity of primary and social
production, which is to say, their relative importance, broadly construed, for
human well-being in a given context.

There are many episodes in history where human settlements transitioned from
dispersed to aggregated, and presumably this is a signal of a change in the
relative value of social vs. primary production for the residents. This can
occur for a variety of reasons, including an increase in the social cost of not
being aggregated (e.g., due to warfare); a decrease in transport costs for
staples (due to animal traction, wheeled vehicles, roads, etc.); an increase in
land productivity which changes the balance of costs of transport for primary
vs. social production; or an increase in the contribution of exchange to
household incomes, which can derive from increases in community size and
density. In addition, while one can conceptualize transitions between dispersed
and aggregated forms of settlement in general systems terms, in real human
communities a variety of political or ideological factors can keep people from
adopting an energetically balanced form of settlement given prevailing
conditions. So, there are opportunities for archaeologists to study deviations
from general equilibrium conditions in addition to factors that shift the
equilibrium in a given context.

For all this work, however, there is a fundamental issue: the identification of
community territories. It is relatively straightforward to define the boundaries
of aggregated settlements by finding the extent of built space and/or
artifactual remains. However, to define the community boundary one needs to know
the extent of agricultural land used by the residents, and this is often
difficult to determine from the village remains themselves [@varien1999]. In
contrast, it is often quite difficult to determine the boundaries of dispersed
communities due to gaps in survey coverage or relatively consistent
distributions of farmsteads, but once one has determined the community territory
it is straightforward to define the associated agricultural land.

In this paper, we present a method we developed to solve such problems, which
enables us to compare the properties and resilience of dispersed vs. aggregated
communities on the same landscapes. We draw on three extensive compilations of
archaeological data from different portions of the ancestral Pueblo region of
the US Southwest, using the locations of what we call community centers as
tethers for defining territories that we infer were used by social communities
over extended periods of time [@glowacki_ortman2012; @varien_etal2007]. In some
regions these are sites containing more households than could plausibly be
related by reckoned kinship or some form of civic architecture. The key feature
of these centers is that they represent locations against which travel time
distances to other sites can be compared.

While not strictly required by our algorithm, we make the simplifying assumption
that once established, community territories were relatively fixed in space for
the duration of their associated sites. This allows us to use the distribution
of all recorded residences, regardless of their periods of occupation, in
defining community territories. There are several reasons why we believe this is
a reasonable simplification. First, there are strong cross-cultural regularities
in the time individuals spend in daily travel from and to their residences, so
dispersed farming communities tend to be dispersed across the distance
associated with this typical travel time, regardless of the number of households
involved [@marchetti1994]. Second, the archaeological record of our study areas
supports the idea that community territories did not shrink when their
populations aggregated into villages. In some cases, farmsteads were converted
to field houses, with some building materials being reused elsewhere and
evidence for continued limited activity use [@varien1999; @varien1999a;
@varien2002]. In other cases, grid gardens were constructed directly within
older rubble mounds by residents of newer, aggregated villages nearby
[@gauthier_herhahn2005]. Third, an analysis that allowed community territories
to vary by time step would reduce variation in the density of the resulting
units, thus washing out variation in one of the most important social properties
one would ideally want to examine. Fourth, in areas where there is strong survey
coverage, dispersed residences form clusters, often but not always centered on a
few larger settlements, and this clustering is apparent even when the dispersed
residences are plotted by time step [@schachner2012; @varien2002]. Fifth, the
effects of population growth for dispersed farming communities are not parallel
to the effects of population expansion in aggregated settlements. In the latter,
the area of the settlement must grow somewhat because the residential density is
already relatively high. But in dispersed settlements, there is often unused
land between existing farmsteads that can be settled. This, combined with the
strong cross-cultural regularity in typical commute times, leads to much greater
consistency in the spatial extent of residents in a dispersed community than in
the case of an aggregated community. Finally, given the inherent incompleteness
of the archaeological record, it seems most appropriate to use all remains to
define what is effectively the maximum extent of community territories. While it
was undoubtedly true that individual community territories expanded or
contracted over time, the assumption of constant area seems more reasonable than
the assumption of constant density.

# Data

The datasets we apply our algorithm to come from three regions in the US
Southwest: Central Mesa Verde (CMV), the Northern Rio Grande (NRG), and the
Cibola or Zuni region (CIB) (see @fig-labeled-overview). The CMV and NRG
datasets derive from Phase II of the Village Ecodynamics Project [@ortman2016;
@schwindt_etal2016]. The CIB data derive from data syntheses produced by
Schachner and Peeples [@peeples2018; @peeples_schachner2012; @schachner2012].
All of these data are now included in cyberSW [@mills_etal2020]. The
distribution of farms and centers is shown in @fig-overview. For summaries of
the archaeological and environmental context in each region, see @tbl-regions.

::: {#fig-labeled-overview fig-align="center"}
![](../figures/labeled-overview-map.png){width="5.2in"}

Overview map showing the locations of the three study areas: Central Mesa Verde,
Northern Rio Grande, and Cibola. Prominent towns, administrative units, and
landforms are also labeled to help orient the reader.
:::

The CMV area encompasses the southwestern corner of Colorado, from Mesa Verde
National Park in the southeast to the Dolores River in the north, and the Utah
state line in the west. The elevation ranges from roughly 1,399.8 m to 3,032.8
m, with an average elevation of 1,999.0 m ($\sigma$ = 276.5 m). SKOPE
[@bocinsky_etal2022] estimates suggest annual precipitation in the CMV study
area has averaged around 39.5 cm ($\sigma$ = 9.1 cm) over the last 2,000 years,
and it puts summer growing degree days for maize around 2,390.1 °F ($\sigma$ =
97.9 °F). Across most areas in the region, it is believed that direct
precipitation was sufficiently high to support dry farming in most years
[@bocinsky_kohler2014; @bocinsky_etal2016].

The NRG study area in north central New Mexico includes basically all of the Rio
Grande valley between the Jemez Mountains to the west and the southernmost
mountains of the Sangre De Cristos to the east, with the Rio Grande itself
running from the northeast down to the western edge of the valley at the foot of
the Pajarito Plateau around Bandelier National Monument. The elevation ranges
from roughly 1,587.9 m to 3,836.2 m, with an average elevation of 2,220.0 m
($\sigma$ = 371.3 m). SKOPE estimates the average annual precipitation in the
NRG over the last 2,000 years to be around 40.6 cm ($\sigma$ = 8.2 cm), and it
puts summer growing degree days for maize around 2,154.4 °F ($\sigma$ = 102.1
°F). It is argued that direct rainfall at Bandelier NM made dry farming possible
there, but in the valley farming required floodwater irrigation along
watercourses [@duwe_anschuetz2013; @bocinsky_kohler2014; @bocinsky_etal2016;
@ortman2016a].

The CIB region straddles the southern slope of the Colorado Plateau along the
border between Arizona and New Mexico, from El Morro National Monument in the
east to the centrally located Zuni Pueblo to the confluence of Jaralosa Draw and
the Zuni River in the southwest. Although this is a rugged terrain by any
reasonable standard, it is notably less so than the other study areas. The
elevation ranges from roughly 1,771.3 m to 2,777.1 m, with an average elevation
of 2,145.9 m ($\sigma$ = 160.0 m). SKOPE estimates suggest annual precipitation
in the CIB area over the last 2,000 years has averaged around 34.3 cm ($\sigma$
= 7.6 cm), and it puts summer growing degree days for maize around 2,317.3 °F
($\sigma$ = 93.8 °F). However, no area in the region receives sufficient direct
rainfall to support dry farming, so some form of water management was required
[@kintigh1985; @muenchrath_etal2002].

::: {#tbl-regions}
```{=latex}

\begin{singlespace}
\small
\begin{tabular}{lllll}
    & 
\textbf{Area [km2]} & 
\textbf{Dates [CE]} & 
\textbf{Archaeology} & 
\textbf{Environment} 
  \\ 
  \hline  
  \noalign{\vskip2mm}  
CMV & 
4,566.4 & 
725-1280 & 
\begin{tabular}[c]{@{}r@{}}
  Rooms: \hfill 68,940\\ 
  Farms: \hfill 5,272\\ 
  Centers: \hfill 170
\end{tabular} & 
\begin{tabular}[r]{@{}r@{}}
  Precipitation: \hfill 39.5 cm ($\sigma$: 9.1)\\ 
  Maize GDD: \hfill 2,390.1 $^{\circ}$F ($\sigma$: 97.9)\\ 
  Elevation: \hfill 1,999.0 m ($\sigma$: 276.5)
\end{tabular} 
  \\[5mm]  
  \hline  
  \noalign{\vskip2mm}  
NRG & 
6,958.0 & 
900-1550 & 
\begin{tabular}[c]{@{}r@{}}
  Rooms: \hfill 46,226\\ 
  Farms: \hfill 2,316\\ 
  Centers: \hfill 181
\end{tabular} & 
\begin{tabular}[r]{@{}r@{}}
  Precipitation: \hfill 40.6 cm ($\sigma$: 8.2)\\ 
  Maize GDD: \hfill 2,154.4 $^{\circ}$F ($\sigma$: 102.1)\\ 
  Elevation: \hfill 2,222.0 m ($\sigma$: 371.3)
\end{tabular} 
  \\[5mm]  
  \hline  
  \noalign{\vskip2mm}  
CIB & 
7,424.1 & 
700-1540 & 
\begin{tabular}[c]{@{}r@{}}
  Rooms: \hfill 26,818\\ 
  Farms: \hfill 713\\ 
  Centers: \hfill 77
\end{tabular} & 
\begin{tabular}[r]{@{}r@{}}
  Precipitation: \hfill 34.3 cm ($\sigma$: 7.6)\\ 
  Maize GDD: \hfill 2,317.3 $^{\circ}$F ($\sigma$: 93.8)\\ 
  Elevation: \hfill 2,145.9 m ($\sigma$: 160.0)
\end{tabular} 
  \\[5mm] 
  \hline
  \noalign{\vskip2mm}
\end{tabular}
\end{singlespace}
```
Regional Summaries
:::

# Methods

Our clustering algorithm draws inspiration from the procedure used by the United
States Office of Management and Budget to define Core Based Statistical Areas
(CBSA) [@usomb2021]. We start by defining a core area, in this case a community
center location. We then merge core areas into larger agglomerations based on
the proportion of their populations that they share. Our method may be described
as a guided density-based clustering algorithm in that it does not rely on a
random set of points when initialized. Instead, the selection of core areas is
guided by archaeological data and regional expertise. We think this is a key
argument in its favor.

From an archaeological perspective, one can think of our algorithm as combining
insights from @varien1999 and @reese_etal2019, specifically Varien's idea of a
community catchment, which is an isochrone defined by a uniform commute time in
all directions from a community center, and Reese's suggestion that a spatial or
geographic community can be identified by grouping individuals in terms of the
commute distances between them. Our way of combining these ideas is to associate
farm sites with specific community centers in terms of their commute distances
and then draw the community catchment around those associated farms. We also
share with them and others [@lipe_hegmon1989; @peterson_drennan2005;
@murdock1949] a focus on "geographic" as opposed to "purely social" communities;
that is, communities constrained in space and comprised of individuals that can
be expected to interact on a regular, even daily, basis.

An important difference for our algorithm is that it incorporates information
about the total mass or population size around each center, so it is at least
conceptually similar to the classic Xtent model proposed by [@renfrew_level1979;
see also @ducke_kroefges2008], as well as the more widely known gravity model
[@isard1954]. In broad outline, it has the following steps:

1.  Identify community centers
2.  Join farms to their nearest community center
3.  Exclude farms beyond a commute time threshold from all centers
    ($D{\text -}max$)
4.  Join centers based on their overlapping populations ($P$, $D{\text -}join$)
5.  Draw smallest concave hull encompassing all farms, centers, and paths

\noindent After step 4, we also apply a filter to remove communities that have
less than a minimal number of dispersed farmsteads, specifically four. Although
this is a somewhat arbitrary threshold, we note that three vertices are the bare
minimum required to define a polygon and measure its area. By removing small
communities, we also minimize potential underestimates of total arable land
area. In addition, our larger ambition is to understand processes occurring
within dispersed farming communities, not necessarily to define all the
dispersed communities that once existed in an area. For community centers with
few or no farms, there is no meaningful dispersed farming community to examine,
regardless of whether this is due to limited survey coverage or actual past
behavior, hence our desire to exclude them.

The primary tuning parameters in our algorithm are $P$, the proportion of a
community population used to join communities; $D{\text -}join$, the commute
time required to calculate P; and $D{\text -}max$, the commute time used to
exclude distant farms. While values for these variables were carefully chosen
based on theory and empirical research, we recognize that they are, at the end
of the day, arbitrary selections, so we include some limited sensitivity
analysis to show how changes to these parameters affect the algorithm (see
@fig-sensitivity). This involves changing the value of a focal parameter while
holding the other parameters at their default values, which we describe below.

::: {#fig-overview fig-align="center"}
![](../figures/overview-maps.png){width="6.5in"}

Overview map showing the locations of farmsteads (black dots) and community
centers (red triangles) in Central Mesa Verde, Northern Rio Grande, and Cibola.
:::

## Identify community centers

**Rule:** Any site, whether residential or otherwise, with known and persistent
interaction between unrelated individuals should count as a community center.

Why persistent interactions? Intuitively, a community center is a place where
individuals from different households cross paths with the intention to interact
[@glowacki_ortman2012; @varien1999; @berrey_etal2021]. A simple way to
demonstrate that intention is to show that individuals would still be willing to
pay some non-negligible cost to visit those centers under various counterfactual
changes to local conditions. This serves to distinguish them from merely
accidental exchanges or chance encounters. Of course, there can be some
path-dependence to this process, with locations of ephemeral or one-off
encounters coming to be locations of persistent interaction. The point is that
the location will come to be a community center so long as individuals continue
to visit that location despite the underlying variability in their
circumstances.

Why unrelated individuals? The simple answer is that a community is not a
family. As a unit of social organization, it does share something in common with
a family, namely, a set of shared interests; but unlike in a family, those
shared interests form only a small part of a much larger set of potentially
conflicting goals and desires. As a consequence, communities are more
susceptible to problems of coordination or collective action [@smith2003;
@smith2010]. On the other hand, the mere existence of a community center implies
some level of coordination - even if that coordination is just about the
location of the community center itself! So, community members share less in
common than families, but more in common than a purely random sample of the
larger population.

We can summarize these ideas by noting that persistent interactions at a
location make that location a *center*, and persistent interactions between
unrelated individuals at that location make it a *community* center
[@varien_potter2008; @gilpin2003; @stone2016; @peterson_drennan2005]. To make
these considerations more useful for defining community boundaries, however, we
need to operationalize them, to make them applicable to specific archaeological
contexts, and that requires a deeper understanding of the archaeology in those
contexts. For given differences in population density and settlement patterning,
the frequency and intensity of interaction between individuals at centers is
likely to vary by region, so it is reasonable to expect that community centers
themselves (at least in so far as they play the role of centers) should vary in
their size and composition across regions, too.

For the two regions represented by the VEP II dataset, we rely on two different
but related sets of criteria to identify community centers. In the Central Mesa
Verde area, a site is considered a community center if it includes civic
architecture, specifically a great kiva or great house, or eight or more pit
structures (small kivas) representing eight or more households. In the Northern
Rio Grande, a site counts as a community center if it is classified as a village
(a site with 50 or more rooms) or town (a site with 500 or more rooms). In the
Cibola region, a community center includes at least 50 rooms or civic
architecture. The large number of households in these regions are thought to be
too large to be plausibly related by reckoned kinship.

## Join farms to their nearest community center

**Rule:** For each farm, assign it to the community of its nearest community
center.

Following @reese_etal2019, we define nearness or proximity in terms of commute
time rather than linear geographic distance. This is done for the obvious reason
that the primary opportunity cost associated with daily pedestrian movement is
the time spent walking, during which other productive activities cannot take
place. There is also a much stronger cross-cultural regularity in the length of
time individuals devote to commuting than the distance covered, as the latter is
a function of the speed of movement, which is influenced by transport technology
[@marchetti1994]. As a result, both the frequency and intensity of interaction
at community centers should decay with commute time. The longer it takes to get
to a community center, the less frequent and less intense the interactions at
that center should be [@varien1999; @peterson_drennan2005].

While different constraints will naturally arise for different modes of
transportation, leading to differences in both the routes taken and the speed
and distance covered, for the populations considered here the only available
mode of transportation was pedestrian. The biggest obstacle to interaction was
thus the landscape itself, in particular its topography - intuitively, the
steeper the uphill slope, the slower the hiking speed, the longer the commute
time. So, all else being equal, individuals at a farm that is equidistant (in
terms of geographic distance) between two community centers should prefer
traveling to the one separated by less rugged terrain.

Commute time is a critical variable in many places in our algorithm, but as a
first approximation, we incorporate the idea by simply joining farms to the
community of their nearest center, meaning the center with the shortest commute
time from the farm. For details of how we estimate commute times over costly
terrains, see "Least-cost path analysis," below.

## Exclude distant farms

**Rule:** Exclude all farms farther than commute time, $D{\text -}max$, from
their nearest center.

The commute time threshold, $D{\text -}max$, defines an isochrone around each
community center. All farm sites that fall outside that isochrone are considered
outliers and dropped from the analysis. Doing this helps to minimize potential
overestimates of total arable land area for each community. This is owing to the
fact that distant farms on average add more area to the community polygon than
nearer farms. Here, we have defined $D{\text -}max$ as one hour. This is about
twice the median daily commute time observed in cross-cultural studies and it
includes roughly 95% of the farmsteads in each region (see @fig-commutes).

## Join centers with overlapping neighbors

**Rule:** For any two centers $c$ and $d$ with populations $N_{c} < N_{d}$, if
$P \cdot N_{c}$ is within distance $D{\text -}join$ of $d$, then $c$ is part of
the same community as $d$.

The larger and more dense farming populations become, the harder it gets to tell
them apart and, more importantly, the harder it gets to justify keeping them
apart, so we need to articulate a rule that specifies when two communities
should be merged into one. For CBSAs, that decision hinges on a distinction
between central and outlying counties, or counties where some fraction of the
population live in an urban core and counties where some fraction commute to and
from a central county on a regular basis. Basically, if the central counties of
one CBSA qualify as the outlying counties of another CBSA, then the two CBSAs
are merged. The intuition here is that when rates of interaction and exchange
between CBSAs are roughly equal to rates within a CBSA, the CBSAs should be
merged [@usomb2021].

We carry that intuition over to our merger rule for dispersed maize farming
communities, albeit with some modification to account for the fact that we are
working with point locations rather than administrative boundaries. We assume
that if some proportion, $P$, of the population related to one community center
lives within a certain commute time, $D{\text -}join$, of another community
center, that those two community centers are effectively parts of the same
dispersed community. We do this in the direction of letting larger communities
absorb smaller ones.

Here we operationalize the population size $N$ of center $c$ using the
well-established rule of thumb in US Southwest archaeology that each surface
room in a settlement represents a single person, leading to a one to one
conversion from rooms to residents [@duwe_etal2016; @kintigh1985; @lekson1989;
@lipe1989; @ortman2016]. In this case, we use the catchment room count, meaning
the number of rooms within $D{\text -}join$ of $c$, defined as:

$$N_{c} = \sum_{i=1}^{S} R_{i} \cdot I(t_{ic} \leq D{\text -}join)$$

\noindent for all sites $S$ (including both farms and centers), with $R_{i}$
being the room count at site $i$, $t_{ic}$ the travel time from $i$ to $c$,
$D{\text -}join$ the threshold travel time to a center, and $I$ the indicator
function that is 1 if $t \leq D{\text -}join$ and 0 otherwise. Note that $S$
includes $c$ and that $t_{cc} = 0$ (the travel time from a center to the same
center is precisely 0), so $R_{c}$ is always included in $N_{c}$.

The catchment room count within $D{\text -}join$ of two centers $c$ and $d$ is
then

$$N_{cd} = \sum_{i=1}^{S} R_{i} \cdot I(t_{ic} \leq D{\text -}join) \cdot I(t_{id} \leq D{\text -}join)$$

\noindent and

$$P_{cd} = N_{cd}/N_{c}$$

\noindent so we could also state our join rule as: if $P_{cd} \geq \alpha$ for
some critical threshold $\alpha$, then $c$ is part of the same community as $d$.
This promotes joins when $c$ and $d$ are themselves in close proximity to each
other, as the farther apart they are, the smaller the number of sites that
overlap within $D{\text -}join$. It also tends to encourage joining two
communities when a substantial proportion of the population lives between the
two centers. In either case, the result is what @varien1999 sometimes refers to
as a "macro-community" or "multi-community cluster," meaning a community
organized around multiple community centers.

The rule was partly inspired by the density-based clustering algorithm known as
DBSCAN [@ester_etal1996], as we rely on a moving window (an isochrone, in this
case) set around each center, and look for some measure of density to guide join
decisions. As noted above, however, our method does not rely on a random
initialization of points, but rather targets community centers. It is also
hierarchical in the sense that it exploits additional structure in our data,
namely the distinction between community centers and farm sites.

An important obstacle to estimating population is that room counts cannot always
be directly measured. For many sites in the CMV, pit structures or small kivas
are more reliably identified than surface rooms. Fortunately, each pit structure
is typically the central roofed space of a unit pueblo residence that includes
an average of six surface rooms [@adler1990], so we simply multiply the number
of pit structures by six to obtain room estimates for each site. CIB and NRG
sites all have room estimates, so no additional assumptions are made about them.

For this analysis, we specify $D{\text -}join$ to be one half hour and $P$ to be
0.8. We acknowledge that these values are somewhat arbitrary. In their defense,
they set an extremely high bar, so joining centers that satisfy the rule seems
reasonable.

## Draw community boundary

**Rule:** For each community, draw a polygon that encompasses all farms,
community centers, and commute paths using a concave hull.

So far, we have only grouped site points into clusters or communities. The goal,
however, is to define their spatial extent. That means, for each set of site
points, we need some way of drawing a polygon that encompasses all of them. This
can be done in a number of ways. The simplest strategy would be to find the
centroid of the set of points and draw a circle with the smallest radius that
still includes them all. This would be utterly arbitrary, however, and frankly
unrealistic, not representing any sort of meaningful boundary given the
landscape. An alternative would be to use the convex hull of the set of points,
but this tends to exaggerate the total area of the community, especially when
the distribution of points is suggestive of a concave shape, like the letter
'C'. A concave hull would handle peculiar shapes, but it would also introduce
the opposite problem, restricting the area of the community to an unreasonable
degree. Our somewhat brute force way of threading this needle is to identify
paths with shortest commute times between the outer most farms in a community,
extract the vertices from those paths, and then incorporate those into the set
of points used to define the concave hull. The result is a polygon that is
concave as well in the dimension of time, meaning the paths with shortest
commute times between any two farms never leave the community.

In some cases, this may lead to overlapping community boundaries, though the
amount of overlap should be minimal given the way that we cluster points,
particularly when there are many densely distributed points. To some extent, of
course, this sort of overlap is both unavoidable and expected as even in the
modern world there are disagreements about the exact locations of legal and
political boundaries.

As a final gloss on community boundaries, we also dilate each concave hull,
briefly expanding it with a positive buffer, then shrinking it with a negative
buffer having a slightly smaller size. This should have only a minimal effect on
the total area of each polygon and also serves to smooth out noise along the
edges.

## Least-cost path analysis

The key to our path analysis is to take information about the terrain in our
study regions and convert it first into meaningful estimates of hiking speed and
then into meaningful estimates of travel time. Fortunately, attempts to model
hiking speed across a range of slopes have advanced considerably in recent
years, most notably in @campbell_etal2019. Drawing on a large, crowd-sourced
fitness tracker data set consisting of approximately 420,000 individual hikes,
jogs, and trail runs from nearly 30,000 individuals in Salt Lake City, UT,
@campbell_etal2019 offer what is easily the most empirically robust estimate of
hiking speed to date. Our application of Campbell's hiking function relies on
coefficients from the fifth percentile of individuals in their sample, using the
Lorentz curve as recommended by Campbell, and assumes that everyone in the areas
of our analysis were largely equal in their walking speed and endurance. This
makes their walking speeds comparable to, but not the same as, what would be
estimated by Tobler's hiking function [@tobler1993].

To calculate travel or commute times, we first download a 1 Arc-second digital
elevation model (DEM) for each study area from the @usgs2023 3D Elevation
Program and convert it into a graph with each node representing a grid cell and
each edge representing a straight line between adjacent grid cells in the Moore
neighborhood (the eight adjacent grid cells). The degree slope of each edge is
calculated, and edges with slope estimates greater than or equal to 45 degrees
are removed. This prevents the path analysis from assuming that a hypothetical
hiker would, for instance, scale steep slopes or walk off a cliff. The remaining
slope estimates are then fed to Campbell's hiking function to derive estimates
of hiking speed along each edge. The inverse of the hiking speed (the pace) is
then multiplied by the distance along that slope to estimate travel time.

Next, we associate each farm and community center with the grid cell that
contains its centroid. In effect, we treat all the sites falling into a grid
cell as a single site. We then apply Dijsktra's algorithm to the entire set of
nodes with associated sites. This gives us two datasets: (i) a dense distance
matrix with least cost travel-times from all farms and community centers to all
farms and community centers and (ii) a set of linestring geometries whose
vertices are the grid cells traversed on shortest paths. For simplicity, we
average travel times to and from each origin and destination point, thus making
the distance matrix symmetric. All of the clustering steps in our algorithm are
implemented as operations on this matrix.

Importantly, the original resolution of the gridded data is \~30 m, but we
aggregate grid cells by a factor of 3, taking the mean in all cases, and making
them closer to \~90 m in resolution. Aggregation is a familiar speed-up
technique for least-cost analysis on large grids containing tens of millions of
cells, though it is not without its risks [@doyle_etal2012;
@cushman_landguth2010; @etherington2016], so we also conduct some limited
sensitivity analysis with varying grid resolutions, which are reported below
(see also @fig-sensitivity).

All analyses, including the cost-distance modeling and our implementation of the
algorithm, are conducted in the R programming language and environment
[@rcoreteam2023]. For more details, please see Supplementary Materials, which
include all code necessary to reproduce these analyses and the figures in the
text.

# Results

The goals of this analysis are (i) to identify dispersed farming communities
based on the distribution of farmsteads and their proximity to community centers
and then (ii) to regionalize those communities by giving them meaningful
boundaries. This was done by developing a simple algorithm inspired by the
concept of a CBSA combined with previous work by @varien1999 and
@reese_etal2019. Using parameter values $D{\text -}join$ = 30 minutes,
$D{\text -}max$ = 1 hour, and $P$ = 0.8, our algorithm identified 104
communities in CMV, 90 communities in NRG, and 45 communities in CIB (see
@fig-communities). The number of communities removed because they included an
insufficient number of sites was 4 for CMV, 35 for NRG, and 23 for CIB, leaving
100, 55, and 22 communities with definable spatial extents.

The mean areas of these communities are 8.68 km$^2$ in CMV, 6.06 km$^2$ in NRG,
and 7.19 km$^2$ in CIB; the mean room counts (for all community centers and
farms) are 647.76 in CMV, 593.49 in NRG, and 766.68 in CIB; and the mean room
densities are 99.69 rooms per km$^2$ in CMV, 219.77 rooms per km$^2$ in NRG, and
238.76 rooms per km$^2$ in CIB. We emphasize that the means in several cases
deviate to a considerable degree from median values (denoted in @tbl-results
with the Greek letter $\eta$), indicating skew in the data and the potential
presence of large outliers. This is illustrated in the boxplots in
@fig-densities, which show the distributions of these and other variables of
interest that can be derived from our analysis and also used to evaluate its
merits (see also @tbl-results).

::: {#fig-communities fig-align="center"}
![](../figures/communities-all.png){width="6.5in"}

Map showing the size, shape, and spatial distribution of community polygons.
Color represents the log density of rooms (the number of rooms per square
kilometer), with lighter colors indicating lower values and darker colors
indicating higher values. Note that densities in each region were re-scaled to
have the same range, so this figure cannot be used to evaluate absolute
differences between regions!
:::

In addition to showing the community polygons, @fig-communities also offers a
simple demographic profile for each region based on the spatial distribution of
all rooms (their count per square kilometer) across communities irrespective of
time period. For visualization purposes, the log of the density is used, as it
helps to normalize the distribution and, in particular, to rein in large
outliers that swamp the variance. It is important to note, too, that values in
each region are re-scaled to be in the same range, so the figure cannot be used
to estimate absolute differences between regions. Still, the relative
differences within each region can be enlightening, but even they must be
interpreted with caution.

Not surprisingly, virtually all of the Mesa Verde NP communities show high
densities. The community that includes Goodman Point and Sand Canyon Pueblo just
north of Ute Mountain also has a relatively high density, as do several of the
communities west of Yellow Jacket and along the Dolores River. In the NRG, the
highest density community is near the center of the region. It juts out from the
Rio Grande at its confluence with the Pojoaque River and includes the San
Ildefonso and Pojoaque puebloes. Another high density community lies in a
shallow north-south canyon at the base of the Sangre de Cristos. Additional high
density communities are observed on the Pajarito Plateau in Bandelier NM. In the
CIB area, the three densest communities can be found in the far eastern edge of
the study area near El Morro NM, with additional dense communities near Ramah
just to the west of El Morro and Manuelito in the northwest.

::: {#tbl-results}
```{=latex}

\begin{singlespace}
\small
\begin{tabular}{lrllllll}
    & 
    & 
\textbf{\begin{tabular}[c]{@{}l@{}}Area\\ (km2)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Rooms\\ (N)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Farms\\ (N)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Centers\\ (N)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Commute\\ (mins)\end{tabular}}  & \textbf{\begin{tabular}[c]{@{}l@{}}Room Density\\ (N/km2)\end{tabular}} 
  \\[2mm]
  \hline
  \noalign{\vskip2mm}
CMV &
\begin{tabular}[c]{@{}l@{}} $\eta$\\ $\mu$\\ $\sigma$ \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 7.1\\ : 8.68\\ : 6.73 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 402\\ : 647.76\\ : 701.36 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 24.5\\ : 44.38\\ : 55.86 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 1\\ : 1.65\\ : 1.37 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 22.35\\ : 23.28\\ : 8.83 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 57.49\\ : 99.69\\ : 97.36 \end{tabular}
  \\[5mm]
  \hline
  \noalign{\vskip2mm}
NRG &
\begin{tabular}[c]{@{}l@{}} $\eta$\\ $\mu$\\ $\sigma$ \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 5.53\\ : 6.06\\ : 4.56 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 479\\ : 593.49\\ : 444.7 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 20\\ : 36.73\\ : 34.96 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 2\\ : 2.49\\ : 2.06 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 19.12\\ : 19.04\\ : 8.77 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 99.43\\ : 219.77\\ : 450.86 \end{tabular}
  \\[5mm]
  \hline
  \noalign{\vskip2mm}
CIB &
\begin{tabular}[c]{@{}l@{}} $\eta$\\ $\mu$\\ $\sigma$ \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 4.37\\ : 7.19\\ : 7.39 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 492\\ : 766.68\\ : 626.09 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 13.5\\ : 24.14\\ : 25.34 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 2\\ : 2\\ : 1.11 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 21.16\\ : 24.74\\ : 12.27 \end{tabular} &
\begin{tabular}[c]{@{}l@{}} : 170.25\\ : 238.76\\ : 270.64 \end{tabular}
  \\[5mm]
  \hline
  \noalign{\vskip2mm}
\end{tabular}
\end{singlespace}
```
$\eta$: median, $\mu$: mean, $\sigma$: standard deviation

Community Summaries
:::

Results of the sensitivity analysis are as expected given the role that the main
tuning parameters play in the algorithm (see @fig-sensitivity). Increasing
$D{\text -}join$ makes it easier to join two communities, so you see growth in
the total area of communities and declines in the total number of resulting
communities as it increases. Increasing $P$ makes it harder to join communities,
so you see fewer and smaller communities as it increases. Increasing
$D{\text -}max$ increases the total area of communities and the average commute
time to centers as it serves to filter out distant farmsteads. Importantly,
finer-grained grid resolutions had no meaningful impact on the results of this
analysis, so the level of aggregation we chose is acceptable. One puzzling
result is that the CIB region shows some volatility in its room densities with
respect to the main tuning parameters. The number of CIB communities is also
largely invariant. We comment on this in the discussion.

# Discussion

Our approach offers several advantages over previous efforts to define dispersed
community territories. First, it handles variation in survey coverage by seeking
to define communities we have evidence for, rather than all communities that may
once have existed. Second, it handles variation in the degree to which surveyors
have lumped or split architectural remains in defining sites [@kintigh2003].
This is a special problem in the CMV. The tradition of site survey recording
within Mesa Verde National Park has generally involved treating each room block
as a separate site, even when there is a continuous artifact scatter between
them. Outside of the park, however, the tradition has been to include all room
blocks within a single artifact scatter as part of a single site. Because our
method incorporates room count estimates, it is robust to these differences in
site definition, enabling properties of the resulting communities to be compared
more directly across these areas. Finally, our method focuses on travel time, a
strong regularity in human affairs, and can be applied to landscapes with
dramatically different topographies. The resulting community territories take
obvious natural topographic barriers to interaction into account, but also allow
territories to cross such boundaries when this is suggested by the data
themselves.

::: {#fig-commutes fig-align="center"}
![](../figures/commute-kl.png){width="4.2in"}

Figure shows the difference between the log density of commute times (in
minutes) from observed farms and random locations to the nearest community
center for each region, along with the location of their 95% quantile (for
observed farms, not random points).
:::

Because our algorithm requires that we calculate the commute time between each
farm and community center, it also allows us to evaluate potential spatial -
strictly, commute time - interactions between them. @fig-commutes tries to get
at this idea by showing the difference between (i) the log probability density
of commute times from observed farms to their nearest centers and (ii) the log
probability density of commute times from random locations to their nearest
centers, along with the 95% quantiles of the observed distributions. This is
analogous to what @reese_etal2019 refer to as the "null difference cost
distance." One can interpret it as showing where commute times differ from what
one might expect were farm sites distributed at random across the landscape.
Values much larger than zero indicate that there are more farms located at those
commute times from centers than you would expect by chance. Conversely, values
less than zero indicate that there are less farms located at those locations
than one might expect by chance. In either case, these differences tell us that
there is positive covariance between farmsteads and nearby centers at commute
times of less than approximately one hour (the value we assigned to
$D{\text -}max$).

The shapes of the density curves are not, however, identical. We attribute this
to three factors: sampling intensity, local topography, and regional settlement
patterning. For instance, the spike in the CIB region at or around the 30 minute
point along with the longer tail (and the larger 95% quantile) is at least
partially owing to the fact that block survey in that area is limited to a few
disconnected locations. The area is also much flatter than in the CMV and NRG
regions, making the background commute distribution much more uniform, at least
within the surveyed areas.

::: {#fig-densities fig-align="center"}
![](../figures/results-boxplot.png){width="5.2in"}

Figure shows for each region the distribution of relevant variables across
communities.
:::

Perhaps more interesting though is the fact that these regions represent
different settlement patterns over different periods of time. In the CMV area,
the pattern of settlement was largely one of growth and dispersal, with local
populations reaching an asymptote after which they spread out into neighboring
areas [@schwindt_etal2016]. Nearly all centers in the CMV area arose within
farming clusters, but they were relatively small compared to those in the other
study areas, and occurred only to a limited degree in the period leading up to
the Great Drought in 1280 CE (and to a lesser degree the period leading up to
and including much of the 10th century), with no large aggregations occurring
afterwards [@varien1999; @glowacki_ortman2012; @ortman2016a]. In the NRG region,
communities had a few dispersed farms early on, more in the period immediately
following 1280 (mostly concentrated in the northern Pajarito Plateau, the area
of Bandelier NM), and then very large aggregations after that [@ortman2016;
@ortman2016a]. As suggested by the relatively small ratio of farms to centers
(see @fig-densities), communities in the CIB region had almost no dispersed
farming for most of the sequence, with only an abbreviated phase of dispersed
settlement that quickly morphed into large aggregations as populations surged in
the period after 1280 [@kintigh1985; @schachner2012; @peeples2018]. As in the
CMV area, aggregations in the NRG and CIB regions sometimes developed within
dispersed farming communities, but unlike in the CMV, many also seem to have
been established after 1280 CE in areas away from where dispersed farms had been
previously located [@schachner2015; @ortman2016a].

These settlement patterns explain why our algorithm dropped 23 communities in
the CIB region, 35 in the NRG, but only 4 in the CMV. Many of the CIB community
centers are simply too isolated from farmsteads for our algorithm to define a
meaningful extent. Settlement patterning probably also explains the volatile
response of the CIB communities to changes in the tuning parameters, as the
small number of farms means small changes are just enough for communities to add
or drop community centers that are excluded by our main analysis.

This is perhaps the biggest limitation of our algorithm. It cannot define
community boundaries when the relationship between community centers and
dispersed farms is not obvious, either because centers are geographically
isolated from farms or because there is an insufficient number of farms to serve
as a bridge between centers. We note, however, that our concept of a community
center does not technically require the presence of an archaeological site.
Community center sites are just a clear signal of the locations of persistent
interaction in the archaeological record.

::: {#fig-sensitivity fig-align="center"}
![](../figures/sensitivity-analysis.png){width="5.5in"}

Figure shows results of sensitivity analysis, focusing on three main parameters
($D{\text -}join$, $D{\text -}max$, and $P$) and approximate grid resolution.
Measures of some key outcomes are also shown, including the number of
communities, their total area, room density, and average commute time from farms
to the nearest community center.
:::

Archaeologists and anthropologists have investigated various dimensions of
community organization and the types of communities they foster. Examples
include biological communities organized around genetic relatedness or physical
similarities [@becker_juengst2017; @kakaliouras2017; @blom2005; @blom2017],
imagined or ideational communities organized around perceived membership and
metaphors [sometimes called *styles of imaginings;* @ortman2011; @anderson1983;
@isbell2000], reproductive communities organized around mating opportunities
[@mahoney_etal2000; @kolb_snead1997], religious communities organized around a
shared sense of the sacred [@bernardini2004; @malville_malville2001], linguistic
communities organized around a shared grammar and vocabulary [@gumperz1968;
@silverstein1998], even sight communities organized around a shared set of
visually striking landmarks [@bernardini_peeples2015]. In fact, there are
probably as many types of communities as there are types of things for people to
care about.

Because the type of community we aim to delineate is a dispersed *farming*
community, we structure our algorithm around the dimensions of food production,
travel time, and social interaction, as these are major constraints on
settlement for maize farmers [@bocinsky_kohler2014; @bocinsky_etal2016;
@vernon_etal2022a; @vernon_etalinpress; @yaworsky_etal2023]. To give just one
example, our algorithm suggests that 50% of CMV communities offer anywhere from
15-50 hectares per farm. For the CIB and NRG regions, that's 25-45 and 15-35
hectares per farm, respectively. This means that they meet the 14 hectare
threshold required for a household to subsist on maize in these regions
[@reese_etal2019; @benson2011; @bocinsky_varien2017]. And that is before taking
into account that the area of the community is held fixed [pace @reese_etal2019]
and all sites are used regardless of time period, so the derived ratios of land
to farms represent minimum possible levels, not the levels actually experienced
by farmers in these communities at any given time.

# Conclusion

In this paper, we develop an algorithm that leverages the difference between
farmsteads and community centers to cluster sites into dispersed farming
communities and define their boundaries. This was inspired by the concept of a
CBSA, along with important contributions from @varien1999 and @reese_etal2019.
Using this algorithm will allow us to capture and compare the changing
properties of both dispersed and aggregated communities using consistent
measurements. For instance, we can use uniform probability density analysis of
pottery assemblages from each site [@ortman2016] to apportion rooms to time
slices and thus estimate changes in population density through time. The ability
to compare these properties for communities of varying scales and densities is
crucial for evaluating their relative costs and benefits for human welfare.

# Acknowledgements

Portions of this research were supported by a grant from the National Science
Foundation (#2213921, to Ortman).

\pagebreak

# References

::: {#refs}
:::