### 2024-06-14 manuscript/jas-reviews-round-two-response.md
final edits

### Response to Reviewer 1

1. Dismissiveness in our response.

    Response: We seem to have communicated poorly in our original response as 
    reviewer one had many important and insightful criticisms of our original 
    submission. In fact, we went to great lengths to incorporate their ideas 
    into the current draft. So, we apologize if there was any misunderstanding
    here.

2. Use the 2022 version of Campbell's hiking function.

    Response: It is indeed an unfortunate oversight on our part that we did not 
    use the most up-to-date version of Campbell's hiking function, especially 
    because one of us (Vernon) worked down the hall from him while he was 
    evidently working on the 2022 paper. At any rate, we thank the reviewer for 
    bringing this to our attention and have incorporated it into the updated 
    analysis.

3. Include a figure to clarify how boundaries are drawn.

    Response: Done! This is now included as an additional figure in the 
    manuscript.

### Response to Reviewer 2

We thank the reviewer for their comments and suggestions for this paper.

### Response to Reviewer 4: Reproducibility Review

1. The Github repository is not accessible.

    Response: The repo was initially set to private, but it has now been made
    public.

2. It would be better to include the uncompiled Quarto document alongside the
   compiled HTML.

    Response: This is now done in two ways: The compiled HTML document now
    has the raw Quarto document embedded in it. This can be accessed by hitting 
    the "</> Code" button at the top of the document and selecting "View 
    Source." The raw content can then be copied and pasted into another a
    text file with a .qmd extension and compiled using Quarto. The HTML document
    also now included a link to the Github repository, which also contains the 
    raw Quarto document.

3. Upload all code and data to another trustworthy repository (meaning, not 
   Github).

    Response: We have uploaded all materials to Zenodo. DOI:

4. Include a README that includes all relevant project metadata.

    Response: This is included in the Github repo and is also uploaded to 
    Zenodo.

5. Include the geopackage "community-centers.gpkg."

    Response: This includes sensitive site location data that we do not have the
    authority to make public.